Digital History in Finland, 1488-1828

### 2017-02-02 inst/examples/20170201-manuscript.Rmd
analyses

>>>>

```{r init2, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
# Read data
dfo <- readRDS("df.Rds")

# Analysis time window
dfo <- dfo %>% filter(publication_year >= 1640 & publication_year <= 1911)

# Calculate title length (word count)
dfo$title_wordcount <- sapply(strsplit(as.character(dfo$title), " "), length)
dfo$title_wordcount[is.na(dfo$title)] <- NA
```

```{r title_length, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
df <- dfo %>% group_by(publication_year) %>%
              summarize(median_title_length = median(title_wordcount),
	      		mean_title_length = mean(title_wordcount),
	      		n = n())

p1 <- ggplot(df, aes(x = publication_year, y = mean_title_length)) +
       geom_point(aes(size = n)) + geom_smooth() + ggtitle("Mean title word count")
p2 <- ggplot(df, aes(x = publication_year, y = median_title_length)) +
       geom_point(aes(size = n)) + geom_smooth() +
       ggtitle("Median title word count")

```{r title_length_by_gatherings, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
top.gatherings <- setdiff(names(top(dfo, "gatherings", 5)), "NA")

df <- dfo %>% filter(gatherings %in% top.gatherings) %>%
              group_by(publication_year, gatherings) %>%
              summarize(mean_title_length = mean(title_wordcount),
	      		n = n())

p <- ggplot(df, aes(x = publication_year, y = mean_title_length,
                     fill = gatherings, color = gatherings)) +
       geom_point(aes(size = n)) +
       geom_smooth() +
       scale_y_log10() +
       ggtitle("Mean title word count (top gatherings)")

```{r title_length_by_lang, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
top.language <- setdiff(names(top(dfo, "language", 11)), "NA")

df <- dfo %>% filter(language %in% top.language) %>%
              group_by(publication_decade, language) %>%
              summarize(mean_title_length = mean(title_wordcount),
	      		n = n())

p <- ggplot(df, aes(x = publication_decade, y = mean_title_length,
                     fill = language, color = language)) +
       geom_point(aes(size = n)) +
       geom_line() +
       #scale_y_log10() +
       ggtitle("Mean title word count (top language)")

### Kungliga / Linkoping

```{r linkoping, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
minyear <- 1640
maxyear <- 1828
myplace <- "Linköping"
cat <- "Kungliga"
df <- df0 %>% filter(publication_place == myplace) %>%
      	      filter(catalog == cat) %>%
              filter(publication_year >= minyear &
	             publication_year <= maxyear)

tab <- unique(df[, c("publication_year", "author", "title", "publisher")])
write.table(tab, sep = "\t", quote = F, row.names = FALSE, file = "output.tables/Linkoping.tab")
```

[Document listing for `r paste(cat, "/", myplace, minyear, "-", maxyear, sep = " ")`](output.tables/Linkoping.tab)

### Title count in Linköping

```{r linkoping2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
p <- plot_timeline(data = df, x = "publication_decade", y = "titlecount") +
       ylab("Title count (n)") +
       ggtitle(paste("Title count: ", cat, "/", myplace))
print(p)
```

### Unique publishers in Linköping by year

Never two publishers seen on the same year.

```{r linkoping3, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
dfs <- df %>% select(publication_year, publisher) %>%
	      filter(!is.na(publisher)) %>%
              unique %>%	      
	      arrange(publication_year) 
kable(dfs)
```

### Title count per capita

The historical population sizes used in this analysis are shown in this [table](https://github.com/rOpenGov/bibliographica/blob/master/inst/extdata/population_sizes_in_cities.csv).

```{r percapita, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
# Read historical population sizes
f <- system.file("inst/extdata/population_sizes_in_cities.csv", package = "bibliographica")
pop <- read.csv(f)
pop <- melt(pop, "Year")
colnames(pop) <- c("publication_decade", "publication_place", "population")
pop <- as_data_frame(pop)
pop$publication_place <- as.character(pop$publication_place)

# Cities with population data
cities <- setdiff(pop$publication_place, "Year")

minyear <- 1640
maxyear <- 1828

# Take cities from Kungliga except Turku
cat <- "Kungliga"
df1 <- df0 %>% filter(publication_place %in% setdiff(cities, "Turku")) %>%
      	      filter(catalog == cat) %>%
              filter(publication_decade %in% pop$publication_decade) %>%
	      select(publication_decade, publication_place) %>%	      
	      group_by(publication_decade, publication_place) %>%
	      summarise(n = n())

# Take cities from Fennica for Turku
cat <- "Kungliga"
df2 <- df0 %>% filter(publication_place %in% "Turku") %>%
      	      filter(catalog == cat) %>%
              filter(publication_decade %in% pop$publication_decade) %>%
	      select(publication_decade, publication_place) %>%	      
	      group_by(publication_decade, publication_place) %>%
	      summarise(n = n())

df <- bind_rows(df1, df2)

# Remove special chars for compatibility (table join fails otherwise)
pop$publication_place <- gsub("Linköping", "Linkoping", pop$publication_place)
df$publication_place <- gsub("Linköping", "Linkoping", df$publication_place)

# Join title count and population size
dfm <- inner_join(pop, df, by = c("publication_decade", "publication_place"))

# Add title count per capita
dfm <- dfm %>% mutate(titles_per_capita = n/population)

p <- ggplot(dfm, aes(x = publication_decade, y = titles_per_capita)) +
       geom_line(aes(color = publication_place), size = 1) +
       geom_point(aes(color = publication_place), size = 2) +
       xlab("Publication decade") +
       ylab("Title count per capita (n)") +
       ggtitle("Title count per capita in selected cities") +
       guides(color = guide_legend(title = "Place", reverse = TRUE))
print(p)

kable(dfm)
```

```{r abstitlecount, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
# Take cities from Kungliga except Turku
cat <- "Kungliga"
df1 <- df0 %>% filter(publication_place %in% setdiff(cities, "Turku")) %>%
      	      filter(catalog == cat) %>%
	      group_by(publication_year, publication_place) %>%
	      summarise(n = n(), paper = sum(paper, na.rm = TRUE))

# Take cities from Fennica for Turku
cat <- "Kungliga"
df2 <- df0 %>% filter(publication_place %in% "Turku") %>%
      	      filter(catalog == cat) %>%
	      group_by(publication_year, publication_place) %>%
	      summarise(n = n(), paper = sum(paper, na.rm = TRUE))

df <- bind_rows(df1, df2)

p <- ggplot(df, aes(x = publication_year, y = n)) +
       geom_line(aes(color = publication_place), size = 1) +
       geom_point(aes(color = publication_place), size = 2) +
       xlab("Publication year") +
       ylab("Title count (n)") +
       ggtitle("Absolute title count in selected cities") +
       guides(color = guide_legend(title = "Place", reverse = TRUE))
print(p)
```

```{r octavo, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
#>2) paper consumption in octavo format books
#>in Stockholm, Turku, Uppsala, Lund, Göteborg and Linköping

dfs <- filter(df, gatherings == "8vo")
p <- ggplot(df, aes(x = publication_year, y = paper)) +
       geom_line(aes(color = publication_place), size = 1) +
       geom_point(aes(color = publication_place), size = 2) +
       xlab("Publication year") +
       ylab("Paper (sheets)") +
       ggtitle("Paper consumption in selected cities") +
       guides(color = guide_legend(title = "Place", reverse = TRUE))
print(p)

#>graph of average title length in words in Stockholm, Turku, Uppsala,
#>Lund, Göteborg and Linköping. Both according to Kungliga, but if
#>possible Turku data from Fennica.

df$title_length <- nchar(df$title)
dfs <- df %>% group_by(publication_year) %>%
              summarise(mean_title_length = mean(title_length, na.rm = TRUE))

p <- ggplot(dfs, aes(x = publication_year, y = mean_title_length)) +
       geom_line(aes(color = publication_place), size = 1) +
       geom_point(aes(color = publication_place), size = 2) +
       xlab("Publication year") +
       ylab("Title length (characters)") +
       ggtitle("Mean title length in selected cities") +
       guides(color = guide_legend(title = "Place", reverse = TRUE))
print(p)
```

## Session info

This document was created with the following versions:

```{r sessioninfo, message=FALSE, warning=FALSE}
sessionInfo()
```

### 2017-02-02 inst/extras/manuscript-krakow.Rmd
analyses

minyear <- 1640
maxyear <- 1828
cat <- "Kungliga"
df <- df0 %>% filter(publication_place %in% cities) %>%
      	      filter(catalog == cat) %>%
              filter(publication_decade %in% pop$publication_decade) %>%
	      select(publication_decade, publication_place) %>%	      
	      group_by(publication_decade, publication_place) %>%
	      summarise(n = n())

@@

p <- ggplot(dfm, aes(x = publication_decade, y = titles_per_capita)) +
       geom_line(aes(color = publication_place), size = 1) +
       geom_point(aes(color = publication_place), size = 2) +
       xlab("Publication decade") +
       ylab("Title count per capita (n)") +
       ggtitle("Title count per capita in selected cities") +
       guides(color = guide_legend(title = "Place", reverse = TRUE))
print(p)       
```

>>>>

minyear <- 1640
maxyear <- 1828

# Take cities from Kungliga except Turku
cat <- "Kungliga"
df1 <- df0 %>% filter(publication_place %in% setdiff(cities, "Turku")) %>%
      	      filter(catalog == cat) %>%
              filter(publication_decade %in% pop$publication_decade) %>%
	      select(publication_decade, publication_place) %>%	      
	      group_by(publication_decade, publication_place) %>%
	      summarise(n = n())

# Take cities from Fennica for Turku
cat <- "Kungliga"
df2 <- df0 %>% filter(publication_place %in% "Turku") %>%
      	      filter(catalog == cat) %>%
              filter(publication_decade %in% pop$publication_decade) %>%
	      select(publication_decade, publication_place) %>%	      
	      group_by(publication_decade, publication_place) %>%
	      summarise(n = n())

df <- bind_rows(df1, df2)

@@

p <- ggplot(dfm, aes(x = publication_decade, y = titles_per_capita)) +
       geom_line(aes(color = publication_place), size = 1) +
       geom_point(aes(color = publication_place), size = 2) +
       xlab("Publication decade") +
       ylab("Title count per capita (n)") +
       ggtitle("Title count per capita in selected cities") +
       guides(color = guide_legend(title = "Place", reverse = TRUE))
print(p)

kable(dfm)
```

```{r abstitlecount, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
# Take cities from Kungliga except Turku
cat <- "Kungliga"
df1 <- df0 %>% filter(publication_place %in% setdiff(cities, "Turku")) %>%
      	      filter(catalog == cat) %>%
	      group_by(publication_year, publication_place) %>%
	      summarise(n = n(), paper = sum(paper, na.rm = TRUE))

# Take cities from Fennica for Turku
cat <- "Kungliga"
df2 <- df0 %>% filter(publication_place %in% "Turku") %>%
      	      filter(catalog == cat) %>%
	      group_by(publication_year, publication_place) %>%
	      summarise(n = n(), paper = sum(paper, na.rm = TRUE))

df <- bind_rows(df1, df2)

p <- ggplot(df, aes(x = publication_year, y = n)) +
       geom_line(aes(color = publication_place), size = 1) +
       geom_point(aes(color = publication_place), size = 2) +
       xlab("Publication year") +
       ylab("Title count (n)") +
       ggtitle("Absolute title count in selected cities") +
       guides(color = guide_legend(title = "Place", reverse = TRUE))
print(p)
```

```{r octavo, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
#>2) paper consumption in octavo format books
#>in Stockholm, Turku, Uppsala, Lund, Göteborg and Linköping

dfs <- filter(df, gatherings == "8vo")
p <- ggplot(df, aes(x = publication_year, y = paper)) +
       geom_line(aes(color = publication_place), size = 1) +
       geom_point(aes(color = publication_place), size = 2) +
       xlab("Publication year") +
       ylab("Paper (sheets)") +
       ggtitle("Paper consumption in selected cities") +
       guides(color = guide_legend(title = "Place", reverse = TRUE))
print(p)

#>graph of average title length in words in Stockholm, Turku, Uppsala,
#>Lund, Göteborg and Linköping. Both according to Kungliga, but if
#>possible Turku data from Fennica.

df$title_length <- nchar(df$title)
dfs <- df %>% group_by(publication_year) %>%
              summarise(mean_title_length = mean(title_length, na.rm = TRUE))

p <- ggplot(dfs, aes(x = publication_year, y = mean_title_length)) +
       geom_line(aes(color = publication_place), size = 1) +
       geom_point(aes(color = publication_place), size = 2) +
       xlab("Publication year") +
       ylab("Title length (characters)") +
       ggtitle("Mean title length in selected cities") +
       guides(color = guide_legend(title = "Place", reverse = TRUE))
print(p)
